---
title: "Lung Cancer Analysis"
author: "Fesobi Oreofeoluwa"
date: "2025-...-..."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Introduction
With what seems to ever increasing factors that are detrimental to health, we always on the look out for causative effects on such factors.
#Methodology
Using R language to run statistical test and plot factors to try to discover any causative factors from this data-set.
##Data Preping
Install Packages to be used in the analysis
```{r}
library(tidyverse)
library(Hmisc)
library(ggplot2)
```
##Load data and explore data 
Data gotten from KAGGLE learn
```{r}
lung_c <- read.csv("C:/Users/USER/Documents/Git_Hub Data/Data/Lung Cancer/Lung_Cancer_Trends_Realistic.csv")

summarise(lung_c)
View(describe(lung_c)) 
```
##Variables 

###Infected 
Patients with lung cancer (of any stage) 
```{r}
lung_c$infect_dummy <- lung_c$Lung_Cancer_Stage != "None"

lung_DeathR <- sum(lung_c$Survival_Status == "Deceased") / length(lung_c$Survival_Status)
```
###Gender
Separated by Gender
```{r}
male <- subset(lung_c, lung_c$Gender== 'Male')
female <- subset(lung_c, lung_c$Gender== 'Female')
```
###Longevity
Those infected with lung cancer (of any stage) but have lived for over 10 years following diagnosis.
```{r}
longev <- subset(lung_c, (2025 - lung_c$Diagnosis_Year) >= 10  & lung_c$Survival_Status == "Alive" &  lung_c$infect_dummy == "TRUE" )

```

##Statistical Testing
T-test between gender
```{r}
lung_c %>% lung_c$infect_dummy <- as.integer(lung_c$infect_dummy) %>% 
   t.test(male$infect_dummy, female$infect_dummy, conf.level = 0.95, alternative = "two.sided")
```

##Plots

###Gender Stats:
```{r pressure, echo=FALSE}
lung_c %>% filter(infect_dummy == "TRUE") %>% 
  ggplot(aes(Gender))+
  geom_bar()+
  labs(title = 'GENDER(INFECTED)'  , y= 'COUNT' , x= 'GENDER')
```

```{r pressure, echo=FALSE}
lung_c %>% filter(infect_dummy == "TRUE") %>% 
  ggplot(aes(Gender))+
  geom_bar()+
  labs(title = 'GENDER(INFECTED)'  , y= 'COUNT' , x= 'GENDER')
```

```{r pressure, echo=FALSE}
lung_c %>% 
  ggplot(aes(Occupation_Exposure, fill = infect_dummy))+
  geom_bar()+
  labs(title = 'OCCUPATIONAL EXPOSURE(INFECTED)'  , y= 'COUNT' , x= 'EXPOSURE')
```

```{r pressure, echo=FALSE}
lung_c %>% 
  ggplot(aes(Diagnosis_Year, fill = infect_dummy))+
  geom_bar()+
  labs(title = 'INFECTION RATE OVER THE YEARS'  , y= 'COUNT' , x= 'YEAR')
```

```{r pressure, echo=FALSE}
lung_c %>% 
  ggplot(aes(Survival_Status, fill = infect_dummy))+
  geom_bar()+
  labs(title = 'SURVIVAL STATS'  , y= 'COUNT' , x= 'SURVIVAL STATUS')
```

### Not Infected
```{r pressure, echo=FALSE}
lung_c %>% filter(infect_dummy == "FALSE") %>%
  ggplot(aes(Survival_Status))+
  geom_bar()+
  labs(title = 'SURVIVAL STATS(NOT INFECTED)'  , y= 'COUNT' , x= 'SURVIVAL STATUS')
```

```{r pressure, echo=FALSE}
lung_c %>% filter(infect_dummy == "FALSE") %>%
  ggplot(aes(BMI))+
  geom_bar()+
  labs(title = "Average BMI(Not infected)")
```

```{r pressure, echo=FALSE}
lung_c %>% filter(infect_dummy == "TRUE") %>%
  ggplot(aes(BMI))+
  geom_bar()
```

```{r pressure, echo=FALSE}
lung_c %>%  filter(infect_dummy == "TRUE") %>%
  ggplot(aes(longevity))+
  geom_bar()+
  labs(title = 'LONGEVITY(ALIVE FOR OVER TEN YEARS FOLLOWING INFECTION)'  , y= 'COUNT' , x= 'LONGEVITY')
```


### Smocking_Status
```{r pressure, echo=FALSE}
lung_c %>% 
  ggplot(aes(Smoking_Status, fill = infect_dummy))+
  geom_bar()+
  labs(title = 'SMOKING STATUS'  , y= 'COUNT' , x= 'SMOKING STATUS')
```

### Secondhand_Smoke
```{r pressure, echo=FALSE}
lung_c %>% ggplot(aes(Secondhand_Smoke_Exposure)) +
  geom_bar() +
  facet_wrap(~lung_c$Lung_Cancer_Stage)+
  labs(title = 'SECOND HAND EXPOSURE'  , y= 'COUNT' , x= 'SMOKE EXPOSURE')
```

###Occupational exposure
```{r pressure, echo=FALSE}
lung_c %>% filter(Lung_Cancer_Stage != "None") %>%
  ggplot(aes(Occupation_Exposure)) +
  geom_bar()+
  labs(title = 'EXPOSURE(INFECTED)'  , y= 'COUNT' , x= 'FACTORS')
```

###Family History
```{r pressure, echo=FALSE}
lung_c %>% filter(Lung_Cancer_Stage != "None") %>%
  ggplot(aes(Family_History, fill = Lung_Cancer_Stage))+
  geom_bar()+
  labs(title = 'FAMILY HISTORY(INFECTED)'  , y= 'cOUNT' , x= 'HISTORY')
```


```{r pressure, echo=FALSE}
lung_c %>%ggplot(aes(Diagnosis_Year, infect_dummy))+
  geom_col()+
  labs(title = 'INFECTION RATE THROUGH THE YEARS'  , y= 'INFECTED COUNT' , x= 'YEARS')
```

###Death rate
```{r pressure, echo=FALSE}
lung_c %>% 
  ggplot(aes(Diagnosis_Year))+
  geom_bar()+
  facet_wrap(~Survival_Status)+
  labs(title = 'INFECTION RATE OVER THE YEARS'  , y= 'COUNT' , x= 'YEAR')
```

###Region 
```{r pressure, echo=FALSE}
lung_c %>% filter(infect_dummy == "TRUE") %>% 
  ggplot(aes(Region))+
  geom_bar()+
  labs(title = "Infection Count by Region(Infected)")
```

```{r pressure, echo=FALSE}
lung_c %>%
  ggplot(aes(Diet_Quality))+
  geom_bar()+
  facet_wrap(~Region)+
  labs(title = "Diet_Quality by Region(Infected)")
```

```{r pressure, echo=FALSE}
lung_c %>%
  ggplot(aes(Smoking_Status))+
  geom_bar()+
  facet_wrap(~Region)+
  labs(title = "Smoking Status by Region(Infected)")
```

```{r pressure, echo=FALSE}
lung_c %>%
  ggplot(aes(Family_History))+
  geom_bar()+
  facet_wrap(~Region)+
  labs(title = "Family History by Region(Infected)")
```

```{r pressure, echo=FALSE}
lung_c %>%
  ggplot(aes(Occupation_Exposure))+
  geom_bar()+
  facet_wrap(~Region)+
  labs(title = "Occupation Exposure by Region(Infected)")
```

```{r pressure, echo=FALSE}
lung_c %>%
  ggplot(aes(Genetic_Markers_Positive))+
  geom_bar()+
  facet_wrap(~Region)+
  labs(title = "Genetic Markers by Region(Infected)")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#Conclusions